Still Missing a Cool Title

### 2006-05-30 talk/dls2006/paper.tex
Filling in numbers, and trying to explain them.

Our tool-chain is capable of translating the Python interpreter of
PyPy, written in RPython, producing right now either ANSI C code as
described before, or LLVM\footnote{the LLVM project is the realisation
of a portable assembler infrastructure, offering both a virtual
machine with JIT capabilities and static compilation. Currently we are
using the latter with its good high-level optimisations for PyPy.}
assembler which is then natively compiled with LLVM tools.

@@

\begin{verbatim}
+------------------------------------+-------------------+-------------------+
|  Interpreter                       | Richards,         | Pystone,          |
|                                    | Time/iteration    | Iterations/second |
+====================================+===================+===================+
|  CPython 2.4.3                     |   789ms    (1.0x) |   40322    (1.0x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c                            |  4269ms    (5.4x) |    7587    (5.3x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-thread                     |  4552ms    (5.8x) |    7122    (5.7x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-stackless                  |  XXX       (6.0x) |    XXX     (6.2x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-gcframework                |  6327ms    (8.0x) |    4960    (8.1x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-stackless-gcframework      |  XXX       (    ) |    XXX     (    ) |
+------------------------------------+-------------------+-------------------+
|  pypy-llvm-c                       |  3797ms    (4.8x) |    7763    (5.2x) |
+------------------------------------+-------------------+-------------------+
|  pypy-llvm-c-prof                  |  2772ms    (3.5x) |   10245    (3.9x) |
+------------------------------------+-------------------+-------------------+
\end{verbatim}

@@

pypy-c-stackless-gcframework
    This variant combines the "gc transformation" step with the
    "stackless transformation" step.  The overhead introduced by the
    stackless feature is balanced with the removal of the overhead of
    pushing and popping roots explicitly on an alternate stack: indeed,
    in this variant it is possible to ask the functions in the current C
    call chain to save their local state and return.  This has the
    side-effect of moving all roots to the heap, where the GC can find
    them.

@@

Other transformations like the "gc" and the "stackless" ones actually
take more time, particuarly when used in combination with each other (we
speculate it is because of the increase in size caused by the previous
transformations).  A translation of pypy-c-stackless, without counting
GCC time, takes 60 minutes; the same for pypy-c-stackless-gcframework
takes XXX minutes.

>>>>

Our tool-chain is capable of translating the Python interpreter of
PyPy, written in RPython, producing right now either ANSI C code as
described before, or LLVM\footnote{The LLVM project is the realisation
of a portable assembler infrastructure, offering both a virtual
machine with JIT capabilities and static compilation. Currently we are
using the latter with its good high-level optimisations for PyPy.}
assembler which is then natively compiled with LLVM tools.

@@

\begin{verbatim}
+------------------------------------+-------------------+-------------------+
|  Interpreter                       | Richards,         | Pystone,          |
|                                    | Time/iteration    | Iterations/second |
+====================================+===================+===================+
|  CPython 2.4.3                     |   789ms    (1.0x) |   40322    (1.0x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c                            |  4269ms    (5.4x) |    7587    (5.3x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-thread                     |  4552ms    (5.8x) |    7122    (5.7x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-stackless                  |  5121ms    (6.5x) |    6060    (6.7x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-gcframework                |  6327ms    (8.0x) |    4960    (8.1x) |
+------------------------------------+-------------------+-------------------+
|  pypy-c-stackless-gcframework      |  8743ms   (11.0x) |    3755   (10.7x) |
+------------------------------------+-------------------+-------------------+
|  pypy-llvm-c                       |  3797ms    (4.8x) |    7763    (5.2x) |
+------------------------------------+-------------------+-------------------+
|  pypy-llvm-c-prof                  |  2772ms    (3.5x) |   10245    (3.9x) |
+------------------------------------+-------------------+-------------------+
\end{verbatim}

@@

pypy-c-stackless-gcframework
    This variant combines the "gc transformation" step with the
    "stackless transformation" step.  The overhead introduced by the
    stackless feature is theoretically balanced with the removal of the
    overhead of pushing and popping roots explicitly on an alternate
    stack: indeed, in this variant it is possible to ask the functions
    in the current C call chain to save their local state and return.
    This has the side-effect of moving all roots to the heap, where the
    GC can find them.  (We hypothesize that the large slowdown is caused
    by the extreme size of the executable in this case - 21MB, compared to
    6MB for the basic pypy-c.  Making it smaller is work in progress.)

@@

Other transformations like the "gc" and the "stackless" ones actually
take more time, particuarly when used in combination with each other (we
speculate it is because of the increase in size caused by the previous
transformations).  A translation of pypy-c-stackless, without counting
GCC time, takes 60 minutes; the same for pypy-c-stackless-gcframework
takes 129 minutes.
